\chapter{Conclusion}
\label{chapter:conclusion}

% ============================================================================
\section{Summary of Findings}
\label{sec:summary}
% ============================================================================

This thesis investigated the context--security trade-off in agentic communication: how can AI agents coordinate across trust boundaries without catastrophic security failures or unsustainable manual overhead? We addressed this question through four research threads.

\textbf{RQ1: What are the failure modes?} Our benchmark revealed five distinct attack classes (\cref{chapter:failure_cases}): direct prompt injection, social engineering, multi-query inference, tool abuse, and memory contamination. Crucially, these classes differ in their amenability to architectural defence. Tool-level filtering eliminates Classes 1 and 4 but is ineffective against inference attacks (Class 3) and memory contamination (Class 5). This taxonomy provides a structured framework for evaluating any future defence mechanism.

\textbf{RQ2: How does a policy file affect utility and security?} Basic tool filtering (Baseline 2) provides substantial security improvements for direct attacks while only modestly reducing utility. However, it leaves significant gaps: continuous state streams cannot be cleanly partitioned, and memory contamination persists. The policy file is a necessary but insufficient building block.

\textbf{RQ3: How should memory be redesigned?} Dual-Track Memory---separating self-state from per-relationship shards at the database level---eliminates cross-relationship leakage without sacrificing context quality within each relationship. The isolation is enforced by metadata filtering in the vector database, making it independent of LLM behaviour. This addresses the ``memory is the weakest link'' finding from our baseline experiments.

\textbf{RQ4: Can adaptive escalation learn social boundaries?} The Intelligent Escalation Protocol demonstrates that precedent-based learning can reduce manual approval overhead over time. By clustering entities and generalising past decisions, the system automates routine approvals while preserving zero-trust guarantees for novel or ambiguous queries.

% ============================================================================
\section{Contributions}
\label{sec:contributions}
% ============================================================================

This thesis makes three contributions:

\begin{enumerate}
    \item \textbf{A dual-measurement benchmark for agentic communication.} We designed the first evaluation framework that measures utility, security, and overhead simultaneously across a trust-tier gradient. The benchmark includes a heartbeat mode for evaluating proactive agent behaviour---an evaluation paradigm that existing capability and security benchmarks do not address. The benchmark can be used to evaluate any agent system's cross-boundary coordination capabilities.

    \item \textbf{Three architectural solutions, each motivated by an empirically observed failure class.} Dual-Track Memory eliminates cross-relationship memory contamination (Class 5). Mountable Context Cells provide capability-based sandboxing that blocks prompt injection and social engineering at the tool level (Classes 1, 2, 4). The Intelligent Escalation Protocol learns social boundaries through precedent clustering, reducing manual overhead while maintaining zero-trust guarantees for novel queries (Class 3 and boundary queries). The incremental ablation design (M0$\to$M1$\to$M2$\to$M3) isolates the contribution of each component.

    \item \textbf{The coordination layer framing.} We articulate the distinction between agent \textit{frameworks} (which provide capabilities to individual agents) and coordination \textit{layers} (which provide trust infrastructure between agents). This framing---analogous to how TCP/IP provides communication infrastructure without dictating what applications run on top---positions secure cross-boundary coordination as an infrastructure problem rather than an application problem.
\end{enumerate}

% ============================================================================
\section{Limitations}
\label{sec:limitations}
% ============================================================================

We identify several limitations that qualify our findings:

\textbf{Scale of evaluation.} Our benchmark uses a curated set of evaluation queries and a fixed number of relationship scenarios. Real-world deployment would involve far more diverse queries, relationships, and edge cases. The benchmark demonstrates feasibility but does not claim exhaustive coverage.

\textbf{Simulated adversaries.} Our adversarial probes, while spanning five attack classes, are designed by the authors. Sophisticated adaptive attackers may discover attack vectors we did not anticipate. The HackAPrompt results \cite{schulhoff2024hackaprompt} suggest that crowd-sourced adversarial testing reveals qualitatively different attacks.

\textbf{No formal verification.} The security guarantees of Mountable Context Cells are architectural, not formally verified. While the MCC design ensures that unauthorised tools and data are absent from the execution environment, we have not proven this property using formal methods (e.g., model checking or information flow analysis).

\textbf{Single-owner assumption.} Our architecture assumes each agent has a single owner. Team-owned agents with hierarchical or consensus-based access control present additional challenges not addressed here.

\textbf{Escalation cold start.} The Intelligent Escalation Protocol requires an initial calibration period during which the owner must manually approve many queries. The overhead during this period may be prohibitive for users who expect immediate autonomous operation.

\textbf{LLM dependence.} Our solutions reduce but do not eliminate dependence on LLM behaviour. The sanitisation agent in the escalation protocol, for instance, is itself an LLM and may be susceptible to adversarial inputs. Defence in depth mitigates but does not eliminate this recursive trust problem.

% ============================================================================
\section{Future Work}
\label{sec:future_work}
% ============================================================================

Several directions extend naturally from this work:

\textbf{Formal verification of context cells.} Applying model checking or information flow analysis \cite{denning1976lattice} to prove that Mountable Context Cells satisfy non-interference properties---no execution trace within a cell can observe data outside the cell's scope.

\textbf{Agent-to-agent protocol.} Our current benchmark evaluates human-to-agent interactions. Extending to fully autonomous agent-to-agent coordination introduces additional challenges: mutual authentication, intent exchange formats, and consensus protocols for multi-party scheduling.

\textbf{Differential privacy for inference attacks.} Multi-query inference (Class 3) is the hardest attack class. Applying differential privacy techniques \cite{dwork2006differential} to bound the cumulative information leakage across multiple queries is a promising direction.

\textbf{Cross-platform generalisation.} Our benchmark is built on a specific agent infrastructure. Evaluating whether the architectural principles (dual-track memory, dynamic sandboxing, precedent escalation) transfer to other agent frameworks (LangChain, AutoGen, CrewAI) would strengthen the generality claim.

\textbf{Longitudinal study.} Measuring how the escalation protocol performs over weeks or months of real use, with genuine (not simulated) relationships and evolving trust dynamics.

% ============================================================================
\section{Concluding Remarks}
\label{sec:concluding_remarks}
% ============================================================================

The coordination overhead that consumes knowledge workers' time will not be solved by more capable individual agents. It requires infrastructure that makes cross-boundary agent communication safe. This thesis demonstrates that architectural enforcement---not prompt-level restriction---is the reliable path to this goal.

The key insight is simple: do not ask the agent to refrain from accessing sensitive data; ensure the sensitive data is not in the room. By physically constructing execution environments that contain only authorised resources, we transform security from a behavioural property (dependent on LLM compliance) into a structural property (enforced by the runtime). This principle---drawn from decades of operating systems research on capability-based security \cite{dennis1966programming}---is, we believe, the foundation upon which practical agent coordination will be built.
