\chapter{Benchmark Design}
\label{chapter:benchmark_design}

Existing agent benchmarks fall into two disconnected categories. \textit{Capability benchmarks}---WebArena \cite{zhou2024webarena}, SWE-bench \cite{jimenez2024swebench}, WebShop \cite{yao2022webshop}---measure task completion but ignore security. \textit{Security benchmarks}---TensorTrust \cite{toyer2024tensortrust}, HackAPrompt \cite{schulhoff2024hackaprompt}---measure attack resistance but ignore the utility cost of the defence. No existing benchmark simultaneously measures both, and none evaluates how these metrics vary across a \textit{gradient of trust relationships}---from a family member who should receive nearly full context, to a stranger who should receive almost none.

This chapter describes the benchmark we built to fill this gap. Our benchmark (i) equips an agent with realistic personal context, (ii) exposes it to external entities across varying trust levels, and (iii) measures utility, security, and manual overhead simultaneously. We describe the agent infrastructure, the communication channel design, the evaluation methodology, and the experiment design.

% ============================================================================
\section{Agent with Personal Context}
\label{sec:agent_context}
% ============================================================================

\subsection{The File and State Abstraction}

Our benchmark agent operates over two types of data, mirroring how a real personal agent would function:

\textbf{Files} ($F$). Persistent knowledge artefacts stored as Markdown notes in a hierarchical folder system. For the benchmark, we populate the agent with realistic personal context: meeting notes, financial summaries, project plans, personal reflections, and research documents. Each file has a unique ID and belongs to one or more folders.

\textbf{States} ($S$). Dynamic, time-varying data sourced from external services via plugins:
\begin{itemize}
    \item \textbf{Calendar}: Events with times, locations, attendees, and descriptions. Access can be granted at three levels: full detail, busy/free only, or none.
    \item \textbf{Email}: Thread summaries with sender, subject, and body. Access can be filtered by keyword (e.g., only emails containing ``stripe'' or ``billing'').
    \item \textbf{Todos}: Task items with priorities, deadlines, and completion status.
\end{itemize}

This abstraction follows the UNIX philosophy---``everything is a file''---extended to include temporal state. The agent loads relevant files and queries relevant states to complete tasks.

\subsection{Implementation}

The agent is implemented as a Next.js API route backed by the Vercel AI SDK for streaming LLM responses with tool use. Tools are defined as typed function signatures that the LLM can invoke:

\begin{itemize}
    \item \texttt{read\_note(\{id\})}: Retrieve the content of a specific note.
    \item \texttt{list\_notes(\{folder\_id?\})}: List available notes, optionally filtered by folder.
    \item \texttt{check\_calendar(\{date\_range\})}: Query calendar events within a date range.
    \item \texttt{check\_calendar\_busy(\{date\_range\})}: Return only busy/free status (no event details).
    \item \texttt{search\_email(\{query\})}: Search email threads by keyword.
    \item \texttt{list\_todos()}: List pending task items.
\end{itemize}

The tool set is not fixed---it is dynamically constructed at runtime based on the access policy for the current interaction. This is the foundation upon which our architectural solutions build.

\subsection{MCP Compatibility}

The tool interface follows Anthropic's Model Context Protocol (MCP) \cite{anthropic2024mcp} specification, making the benchmark infrastructure compatible with any MCP-compliant agent. External agents can connect to the benchmark through the standard MCP tool discovery and invocation protocol, enabling evaluation of third-party agents against our benchmark without modification.

% ============================================================================
\section{Agent Communication Channel}
\label{sec:communication_channel}
% ============================================================================

\subsection{The 1-to-1 Interaction Model}

A critical design choice: our communication model is \textbf{1-to-1 instant messaging}, not group chat. Each external entity $E_i$ interacts with the host agent through an isolated session. There is no shared context between sessions.

This choice is deliberate. Group chat introduces combinatorial complexity in access control (what can $E_i$ see that $E_j$ said?) and makes memory isolation intractable. The 1-to-1 model cleanly maps to our formal problem: each session is a single $(A, E_i)$ interaction governed by one policy.

\subsection{Policy Files}

Each relationship is governed by a \textbf{policy file} that specifies what the agent may access and disclose during that interaction:

\begin{verbatim}
{
  "relationship_id": "investor_a",
  "trust_tier": "formal_business",
  "allowedNoteIds": [123, 456, 789],
  "calendarAccess": "busy_only",
  "emailAccess": "filtered",
  "emailFilter": ["stripe", "billing"],
  "allowedTools": ["read_note", "check_calendar_busy",
                    "search_email"]
}
\end{verbatim}

The policy file is the minimal viable access control mechanism. It defines which notes are visible, what level of calendar access is granted, whether and how email access is filtered, and which tools are available. For Baseline 1 (naive prompting), no policy file is used---the agent has unrestricted access. For Baseline 2, the policy file is enforced through tool filtering.

\subsection{Access Granting via Shared Links}

External entities access the agent through \textbf{agentic links}---unique URLs that embed a token referencing the policy file. When a guest opens the link, the backend retrieves the associated policy and constructs the execution environment accordingly. This mechanism is analogous to how Notion or Google Docs share links with embedded permissions---but instead of granting access to a static document, it grants access to a sandboxed agent.

% ============================================================================
\section{Evaluation Methodology}
\label{sec:evaluation_methodology}
% ============================================================================

\subsection{Dual Measurement: Utility and Security}

Most existing agent benchmarks measure only capability (task completion). Security benchmarks measure only attack resistance. Our benchmark measures \textbf{both simultaneously}, because the trade-off between them is the core research question. A defence that achieves 100\% security by refusing all queries is trivial and useless; a system that achieves 100\% utility by answering everything is maximally insecure. The interesting question---and the one no existing benchmark can answer---is \textit{where the Pareto frontier lies} and how architectural interventions shift it.

For each evaluation scenario, we measure:
\begin{itemize}
    \item \textbf{Utility} ($\mathcal{U}$): The fraction of legitimate queries answered correctly and completely. A query is ``legitimate'' if it falls within the guest's policy scope.
    \item \textbf{Security} ($\mathcal{S}$): The fraction of adversarial probes that are correctly blocked. A probe ``succeeds'' (security failure) if the agent discloses information outside the guest's policy scope.
    \item \textbf{Overhead} ($\mathcal{O}$): The fraction of queries that are escalated to the owner for manual approval.
\end{itemize}

\subsection{Heartbeat Mode}
\label{sec:heartbeat}

Static question-answering benchmarks miss an important class of failures: those that arise from \textbf{autonomous agent behaviour}. Real agents do not passively wait for queries---they monitor state changes and take proactive actions.

We introduce \textbf{heartbeat mode}: instead of static QA, the agent operates autonomously at regular intervals (e.g., every 5 minutes). At each heartbeat, the agent:
\begin{enumerate}
    \item Checks for state changes (new emails, calendar updates, completed todos).
    \item Decides whether any changes are relevant to active relationships.
    \item Takes autonomous actions (send notifications, update shared context, generate summaries).
\end{enumerate}

After $N$ heartbeat rounds, we evaluate: what information was disclosed? Were any proactive notifications inappropriate given the guest's policy? Did the agent act on stale or contaminated memory?

Heartbeat mode is critical because it evaluates the agent as a \textit{persistent, proactive entity}---the mode in which real agents will operate---rather than as a passive responder.

\subsection{Evaluation Queries}

We construct three categories of evaluation queries:

\begin{enumerate}
    \item \textbf{Legitimate queries} (utility measurement): Questions that fall within the guest's policy scope. ``When is the owner free next Tuesday?'' (with calendar access). ``What does the Q1 summary say about revenue?'' (with access to note 123).

    \item \textbf{Adversarial probes} (security measurement): Attacks spanning the five classes defined in \cref{chapter:failure_cases}---direct prompt injection, social engineering, multi-query inference, tool abuse, and memory contamination.

    \item \textbf{Boundary queries} (escalation measurement): Ambiguous queries where the correct response depends on relational context. ``Can you share the latest investor update?'' may be appropriate for a board member but not a casual contact.
\end{enumerate}

% ============================================================================
\section{Experiment Design}
\label{sec:experiment_design}
% ============================================================================

\subsection{Agent Configurations}

We evaluate four configurations, each adding one layer of architectural protection:

\begin{enumerate}
    \item \textbf{Baseline 1: Naive prompting.} The agent has full access to all files, states, and tools. Security relies entirely on the LLM's instruction-following.

    \item \textbf{Baseline 2: Policy file with tool filtering.} The agent's tool set is filtered by the policy file. Unauthorised tools are removed from the function array.

    \item \textbf{Solution A: Baseline 2 + Dual-Track Memory.} Relationship-specific memory shards are enforced at the database level.

    \item \textbf{Solution B: Solution A + Dynamic Sandboxing (MCC).} The full Mountable Context Cell is constructed for each interaction.

    \item \textbf{Solution C: Solution B + Intelligent Escalation.} Precedent-based learning automates approval for routine boundary queries.
\end{enumerate}

\subsection{Trust-Tier Scenarios}

We define three trust tiers representing the spectrum of real-world relationships:

\begin{itemize}
    \item \textbf{Inner Circle} (cofounder, close collaborator): Broad access to files and states. Few restrictions.
    \item \textbf{Formal Business} (investor, client): Access to specific files and filtered states. Calendar busy/free only.
    \item \textbf{Acquaintance} (casual contact, stranger): Minimal access. Public information only.
\end{itemize}

Each trust tier has a corresponding policy file. We run the full evaluation suite at each tier to measure how utility and security vary with trust level.

\subsection{LLM Backends}

To ensure our findings generalise beyond a single model, we evaluate with multiple LLM backends, including both frontier models (GPT-4, Claude) and smaller open-source alternatives. This allows us to distinguish architectural effects from model-specific behaviour.

\subsection{Ablations}

We design targeted ablations to isolate the contribution of each component:
\begin{itemize}
    \item \textbf{Policy file ablation}: Baseline 1 vs.\ Baseline 2 isolates the effect of basic tool filtering.
    \item \textbf{Memory ablation}: Solution A vs.\ Baseline 2 isolates the effect of dual-track memory.
    \item \textbf{Sandboxing ablation}: Solution B vs.\ Solution A isolates the effect of full context cell construction.
    \item \textbf{Escalation ablation}: Solution C vs.\ Solution B isolates the effect of precedent-based escalation.
\end{itemize}
