\begin{abstract}

AI agents can now autonomously browse the web, manage calendars, write code, and orchestrate complex workflows. Yet they remain fundamentally siloed: no infrastructure exists for agents to communicate securely across trust boundaries. This creates a paradox---effective coordination requires deep personal context, but exposing context-rich agents to external parties creates catastrophic security risks. Prompt-level defences have been shown to fail under adaptive attacks, and manual permission management reintroduces the coordination overhead agents were designed to eliminate.

This thesis investigates the \textbf{context--security trade-off} in agentic communication. We make three contributions. First, we design a \textbf{benchmark} that measures both collaborative utility and security simultaneously, including a heartbeat mode for evaluating proactive agent behaviour. Red-teaming reveals five classes of failure: direct prompt injection, social engineering, multi-query inference, tool abuse, and memory contamination.

Second, motivated by these failure cases, we propose three \textbf{architectural solutions}: (i)~\textit{Dual-Track Memory}, which isolates episodic memory per relationship at the database level, preventing cross-contamination; (ii)~\textit{Mountable Context Cells}, capability-based execution sandboxes that physically remove unauthorised tools and data from the agent's context; and (iii)~an \textit{Intelligent Escalation Protocol} that learns social boundaries through precedent clustering, reducing manual approval overhead over time.

Third, we articulate the distinction between agent frameworks and \textbf{coordination layers}---positioning our work as trust infrastructure that is compatible with any agent system, analogous to how TCP/IP enables communication between heterogeneous applications.

We evaluate these solutions on our benchmark with multiple LLM backends, measuring the Pareto frontier between utility, security, and manual overhead. Our results demonstrate that architectural enforcement---not prompt-level restriction---is the reliable path to secure agentic communication.

\end{abstract}
