\chapter{Problem Formulation}
\label{chapter:problem_formulation}

This chapter formalises the \textbf{context--security trade-off} that arises when AI agents communicate across trust boundaries. We define the problem mathematically, state the threat model, and derive the four research questions that guide the remainder of this thesis.

% ============================================================================
\section{The Context--Security Trade-off}
\label{sec:trade_off}
% ============================================================================

\subsection{Formal Problem Statement}

Let $A$ be a host AI agent owned by user $U$, possessing access to:
\begin{itemize}
    \item $F = \{f_1, f_2, \ldots, f_n\}$: a set of \textbf{files} (notes, documents, knowledge artefacts);
    \item $S = \{s_1, s_2, \ldots, s_m\}$: a set of \textbf{dynamic states} (calendar events, email threads, todo items);
    \item $M_{\text{self}}$: \textbf{self-state memory}---core facts about $U$ (preferences, constraints, identity).
\end{itemize}

When $A$ interacts with an external entity $E_i$ (another user's agent, or a human), the system must simultaneously satisfy three requirements:

\begin{enumerate}
    \item \textbf{Maximise utility.} Complete collaborative tasks $T = \{t_1, t_2, \ldots, t_k\}$ (schedule meetings, answer questions, share relevant context) by leveraging appropriate subsets of $F$, $S$, and $M_{\text{self}}$.

    \item \textbf{Enforce security.} Guarantee that $E_i$ can access only the authorised subsets $F_i \subseteq F$ and $S_i \subseteq S$, despite adversarial inputs attempting to extract $F \setminus F_i$ or $S \setminus S_i$.

    \item \textbf{Maintain isolation.} Ensure that memory artefacts from the interaction with $E_i$ do not contaminate interactions with $E_j$ for $i \neq j$.
\end{enumerate}

The trade-off is inherent: increasing the agent's access to context (larger $F_i$, $S_i$) improves utility but increases the attack surface. Restricting access improves security but reduces the agent's ability to complete tasks. Manual management of the $(E_i, f_j, s_k)$ permission triples scales quadratically and reintroduces the coordination overhead agents were meant to eliminate.

\subsection{The Pareto Frontier}

We model the trade-off as a multi-objective optimisation problem. For a given policy configuration $\pi$ and agent architecture $\alpha$, define:

\begin{itemize}
    \item $\mathcal{U}(\pi, \alpha)$: task completion rate across all legitimate queries.
    \item $\mathcal{S}(\pi, \alpha)$: data protection rate across all adversarial probes.
    \item $\mathcal{O}(\pi, \alpha)$: manual overhead---the fraction of queries requiring human intervention.
\end{itemize}

The ideal system achieves $\mathcal{U} = 1$, $\mathcal{S} = 1$, $\mathcal{O} = 0$. In practice, these objectives conflict. The goal of this thesis is to push the \textbf{Pareto frontier}---improving security without proportionally sacrificing utility or increasing overhead.

% ============================================================================
\section{Threat Model}
\label{sec:threat_model}
% ============================================================================

We adopt a \textbf{Byzantine adversary model} where external entities $E_i$ may be actively hostile. Specifically, we assume the adversary may:

\begin{enumerate}
    \item \textbf{Inject adversarial prompts.} Attempt to override the agent's system instructions through direct \cite{greshake2023not} or indirect \cite{zou2023universal} prompt injection.

    \item \textbf{Employ social engineering.} Claim false authority (``The owner gave me access'') or appeal to the agent's helpfulness bias \cite{wei2024jailbroken}.

    \item \textbf{Execute multi-query inference.} Ask a sequence of individually innocuous questions whose answers, when combined, reveal protected information.

    \item \textbf{Abuse tool interfaces.} Invoke tools with out-of-policy arguments or exploit implicit tool capabilities.

    \item \textbf{Exploit memory contamination.} Attempt to retrieve information stored from other relationships through the agent's shared memory.
\end{enumerate}

We do \textit{not} assume compromise of the underlying LLM service, the operating system, or the network transport layer---these are orthogonal security concerns addressed by existing infrastructure. Our threat model focuses on attacks that exploit the \textit{application layer}: the interface between the LLM, its tools, and its memory.

This threat model aligns with Meta's ``Rule of Two'' framework \cite{meta2025ruletwo}: an agent becomes critically vulnerable when it simultaneously processes untrusted inputs, accesses sensitive data, and can take state-changing actions. Our solutions aim to ensure that at most two of these three conditions hold at any time.

% ============================================================================
\section{Limitations of Existing Approaches}
\label{sec:existing_limitations}
% ============================================================================

Before stating our research questions, we summarise why existing defences are insufficient for the problem as formulated:

\textbf{Prompt-level defences} (system instructions, content classifiers, chain-of-thought filtering) improve average-case security but fail under adaptive attacks. Nasr et al.\ \cite{nasr2025attacker} tested twelve published defences and found all could be bypassed with $>$90\% success. Even Anthropic's state-of-the-art model-level hardening leaves a 1\% residual attack success rate \cite{anthropic2025promptinjection}---unacceptable when personal data is at stake.

\textbf{Static access control lists} (file-level permissions) provide strong isolation for static documents but cannot handle continuous state streams (email, calendar) where new content arrives constantly and sensitivity depends on relational context.

\textbf{Manual permission management} offers perfect security but reintroduces the coordination overhead that agents were designed to eliminate. If the owner must approve every cross-boundary query, the agent provides no net productivity gain.

The research community has converged on the conclusion that \textbf{architectural enforcement}---physically constraining what the agent can access, rather than instructing it what not to access---is the reliable path forward \cite{google2025promptinjection, meta2025ruletwo}. Our solutions build on this principle.

% ============================================================================
\section{Research Questions}
\label{sec:research_questions}
% ============================================================================

The problem formulation and the limitations of existing approaches motivate four research questions:

\begin{description}
    \item[RQ1] \textit{What are the failure modes when AI agents communicate across trust boundaries?}

    We seek a systematic taxonomy of how agents fail---not just prompt injection, but social engineering, inference attacks, tool abuse, and memory contamination. Understanding the failure landscape is prerequisite to designing effective defences. (\Cref{chapter:failure_cases}.)

    \item[RQ2] \textit{How does a policy file (access rules) affect both utility and security?}

    Basic tool filtering is the simplest architectural intervention. We measure its effect on both legitimate task completion and adversarial attack success, establishing the baseline from which our solutions improve. (\Cref{chapter:failure_cases,chapter:experiments}.)

    \item[RQ3] \textit{How should agent memory be redesigned for multi-party interactions?}

    Global memory creates cross-relationship leakage. We ask whether a dual-track architecture---separating self-state from per-relationship shards at the database level---can eliminate memory contamination without sacrificing context quality within each relationship. (\Cref{chapter:solution_proposal,chapter:experiments}.)

    \item[RQ4] \textit{Can an adaptive escalation mechanism learn social boundaries and reduce manual overhead?}

    Manual approval for every ambiguous query is unsustainable. We ask whether precedent-based learning---clustering entities and generalising past decisions---can automate approval for routine queries while preserving zero-trust guarantees for novel ones. (\Cref{chapter:solution_proposal,chapter:experiments}.)
\end{description}

These four questions span the full problem space: RQ1 characterises the threat landscape, RQ2 establishes baselines, RQ3 addresses memory isolation, and RQ4 tackles the overhead problem. Together, they trace a path from \textit{diagnosis} (what goes wrong, \cref{chapter:failure_cases}) through \textit{treatment} (architectural solutions, \cref{chapter:solution_proposal}) to \textit{evaluation} (do the solutions shift the Pareto frontier, \cref{chapter:experiments}). Answering these questions requires a benchmark that measures $\mathcal{U}$, $\mathcal{S}$, and $\mathcal{O}$ simultaneously---which we design in \cref{chapter:benchmark_design}.
